---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Working directory set to correct folder
Load data and read file 

```{r Loading data}
library(dplyr)
library(ggplot2)
activity <- read.csv("activity.csv")
str(activity)
summary(activity)
head(activity)
```

Process data into suitable format for analysis
We need to remove missing values. Create another dataset without missing values

```{r Formatiing data}
act.complete <- na.omit(activity)
```

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.
1.Calculate the total number of steps taken per day

```{r stepsperday}
  act.day <- group_by(act.complete,date)
  act.day <- summarize(act.day,steps = sum(steps))
  summary(act.day)
```  


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
